---
title: Feature-based Time Series Forecasting
author: Thiyanga Talagala, \break Rob J Hyndman, George Athanasopoulos, \break Feng Li, Yanfei Kang
date:  "11 July 2019"
fontsize: 12pt
classoption: compress
toc: false
output:
  binb::monash:
    fig_height: 5
    fig_width: 8
    highlight: tango
    incremental: no
    keep_tex: no
    includes:
      in_header: preamble.tex
---


```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE, cache=FALSE, dev.args=list(bg=grey(0.9), pointsize=11))
library(tsfeatures)
library(tidyverse)
library(forecast)
library(Mcomp)
library(pander)
library(seer)
```



## Big picture

\begin{tikzpicture}
    \node<1> (img1) {\centerline{\includegraphics[width=\paperwidth, height=6.3cm]{img/rice1.png}}};
    \node<2> (img2) {\centerline{\includegraphics[width=\paperwidth, height=6.3cm]{img/rice2.png}}};
\end{tikzpicture}

\begin{itemize}[<+->]
\vfill\item What algorithm is likely to perform best?
\vfill\item Algorithm selection problem, John Rice (1976)
\end{itemize}


## Time series features

\fontsize{13}{14}\sf

* Transform a given time series $y=\{y_1, y_2, \cdots, y_n\}$ to a feature vector $F = (f_1(y), f_2(y), \cdots, f_p(y))'$.

\pause

\begincols
\begincol{.48\textwidth}


```{r, warning=FALSE, message=F, echo=FALSE, fig.height=10}
library(grid)
library(gridExtra)
library(ggrepel)
library(png)
library(tsfeatures)
library(ggpubr)

# Extract required series
series_id <- c("N0001", "N0633", "N0625", "N0645","N1912", "N2012")
six_series <- lapply(M3[series_id], function(u){u$x})
p <- lapply(six_series,
function(u) {autoplot(u) + xlab("") + ylab("")}
)
for (i in seq_along(six_series))
p[[i]] <- p[[i]] + ggtitle(series_id[i])+theme(title =element_text(size=30, face='bold'))
grid.arrange(grobs = p, ncol = 2)

```

\endcol 
\begincol{.48\textwidth}

\pause

```{r fig2, fig.height=6.8, fig.width=7, message=F, echo=F, warning=F}
df <- tsfeatures(six_series, c("stl_features")) %>%
  select(trend, seasonal_strength) %>%
  rename(seasonality = seasonal_strength) %>%
  replace_na(list(seasonality = 0))
df$id <- names(six_series)
ggplot(df, aes(x = trend, y = seasonality)) +
  geom_point(size = 5, color = 'red') +
  xlim(0, 1) + ylim(0, 1) +
  coord_fixed() +
  geom_text_repel(
    aes(label = id),
    colour = "black",
    size = 10,
    box.padding = unit(0.5, "lines")
  ) +
  theme(legend.position = "none", text = element_text(size=20))
```

  \endcol
\endcols

## More features

\begin{textblock}{12}(0.1,1.3)
\begin{multicols}{2}
  \begin{itemize}\tightlist
    \item length
    \item strength of seasonality
    \item strength of trend
    \item linearity
    \item curvature
    \item spikiness
    \item stability
    \item lumpiness
    \item spectral entropy
    \item Hurst exponent
    \item nonlinearity
    \item unit root test statistics
    \item parameter estimates of Holt's linear trend method
    \item parameter estimates of Holt-Winters' additive method
    \item ACF and PACF based features - calculated on raw, differenced, seasonally-differenced series and remainder series.
    \end{itemize}
\end{multicols}
\end{textblock}


## Algorithm selection framework

\begin{tikzpicture}
    \node<1> (img1) {\centerline{\includegraphics[width=\paperwidth]{img/f1.png}}};
    \node<2> (img2) {\centerline{\includegraphics[width=\paperwidth]{img/f2.png}}};
    \node<3> (img3) {\centerline{\includegraphics[width=\paperwidth]{img/f3.png}}};
    \node<4> (img4) {\centerline{\includegraphics[width=\paperwidth]{img/f4.png}}};
    \node<5> (img5) {\centerline{\includegraphics[width=\paperwidth]{img/f5.png}}};
    \node<6> (img6) {\centerline{\includegraphics[width=\paperwidth]{img/f6.png}}};
    \node<7> (img7) {\centerline{\includegraphics[width=\paperwidth]{img/f7.png}}};
    \node<8> (img8) {\centerline{\includegraphics[width=\paperwidth]{img/f8.png}}};
    \node<9> (img9) {\centerline{\includegraphics[width=\paperwidth]{img/f9.png}}};
    \node<10> (img10) {\centerline{\includegraphics[width=\paperwidth]{img/f10.png}}};
    \node<11> (img11) {\centerline{\includegraphics[width=\paperwidth]{img/f11.png}}};
    \node<12> (img12) {\centerline{\includegraphics[width=\paperwidth]{img/f12.png}}};
    \node<13> (img13) {\centerline{\includegraphics[width=\paperwidth]{img/f13.png}}};
    \node<14> (img14) {\centerline{\includegraphics[width=\paperwidth]{img/fh14.png}}};
\end{tikzpicture}

## FFORMS: Feature-based FORecast Model Selection

 \centerline{\includegraphics[width=\paperwidth]{img/fformsd1.png}}
 
* \fontsize{8}{8}{two algorithms: \colorbox{violet}{\color{white}{FFORMS}}, FFORMPP}

## \fontsize{12}{12}\bf\sffamily FFORMPP: Feature-based FORecast Model Performance Prediction

 \centerline{\includegraphics[width=\paperwidth]{img/fformppd1.png}}
 
* \fontsize{8}{8}{two algorithms: \colorbox{violet}{\color{white}{FFORMS}}, \colorbox{cyan}{\color{black}{FFORMPP}}}


## `seer` R package

\textbf{Installation}
```r
devtools::install_github("thiyangt/seer")
library(seer)
```
\begin{tikzpicture}[remember picture,overlay]  
  \node [xshift=-1.5cm,yshift=-2cm] at (current page.north east)
    {\includegraphics[width=\textwidth,height=3cm,keepaspectratio]{img/seer.png}};
\end{tikzpicture}

\pause

\textbf{Example dataset}

\colorbox{blue}{\textbf{\color{white}{observed time series - M1 yearly series (181)}}}

```r
library(Mcomp)
yearlym1 <- subset(M1, "yearly")
```
\pause

```{r echo=FALSE, message=FALSE, warning=FALSE, comment=NA, size='tiny'}
library(Mcomp)
yearlym1 <- subset(M1, "yearly")
```



## Input: features

```r
cal_features(yearlym1[1:2], database="M1",
h=6, highfreq=FALSE)

```
\fontsize{8}{10}\sf


```{r echo=FALSE, message=FALSE, warning=FALSE, comment=NA, size='tiny'}
seer::cal_features(yearlym1[1:2], database="M1", h=6, highfreq=FALSE) 
```

## Output: class labels

\fontsize{10}{10}\sf
```r 
seer::fcast_accuracy(tslist=yearlym1[1:2], 
               models= c("arima","ets","rw", "theta", "nn"), 
               database ="M1", cal_MASE, h=6, 
               length_out = 1, 
               fcast_save = TRUE)
```

\fontsize{6}{6}\sf
```{r echo=FALSE, message=FALSE, warning=FALSE, comment=NA, size='tiny'}
seer::fcast_accuracy(tslist=yearlym1[1:2], 
               models= c("arima","ets","rw","theta", "nn"), 
               database ="M1", cal_MASE, h=6, 
               length_out = 1, 
               fcast_save = TRUE)
```

## Training set

```r
prepare_trainingset(accuracy_set = accuracy_m1, 
feature_set = features_m1)$trainingset
```

\fontsize{8}{10}\sf

```{r echo=FALSE, message=FALSE, warning=FALSE, comment=NA, size='tiny'}
data(M1)
yearly_m1 <- subset(M1, "yearly")
accuracy_m1 <- fcast_accuracy(tslist=yearly_m1[1:2], models= c("arima","ets","rw","rwd", "theta", "nn"), database ="M1", cal_MASE, h=6, length_out = 1, fcast_save = TRUE)
features_m1 <- cal_features(yearly_m1[1:2], database="M1", h=6, highfreq = FALSE)

# prepare training set
prepare_trainingset(accuracy_set = accuracy_m1, feature_set = features_m1)$trainingset
```

## FFORMS classifier

\fontsize{12}{12}\sf
```r
rf <- build_rf(training_set = training_set, 
                testset= M3yearly_features,  
                rf_type="ru", ntree=100, seed=1, 
               import=FALSE, mtry = 8)
```


```{r echo=FALSE, message=FALSE, warning=FALSE, comment=NA, size='tiny'}
# steps 3 and 4 applied to yearly series of M1 competition
data(M1)
yearly_m1 <- subset(M1, "yearly")
accuracy_m1 <- fcast_accuracy(tslist=yearly_m1, models= c("arima","ets","rw","rwd", "theta", "nn"), database ="M1", cal_MASE, h=6, length_out = 1, fcast_save = TRUE)
features_m1 <- cal_features(yearly_m1, database="M1", h=6, highfreq = FALSE)
# prepare training set
prep_trainingset <- prepare_trainingset(accuracy_set = accuracy_m1, feature_set = features_m1)
training_set <- prep_trainingset$trainingset
yearly_m3 <- subset(M3, "yearly") # M3 dataset is used as a test set
M3yearly_features <- seer::cal_features(yearly_m3, database="M3", h=6, highfreq = FALSE)
rf <- build_rf(training_set = training_set, testset= M3yearly_features,  
               rf_type="ru", ntree=100, seed=1, 
               import=FALSE, mtry = 8)
```
\fontsize{8}{8}\sf
\colorbox{red}{\textbf{\color{white}{Predictions}}}
\fontsize{8}{8}\sf
```r
head(rf$predictions)
```

```{r, eval=TRUE, size='tiny'}
head(rf$predictions)
```

\colorbox{red}{\textbf{\color{white}{FFORMS classifier}}}


\fontsize{8}{8}\sf
```r
rf$randomforest
```
```{r, eval=TRUE, size='tiny'}
rf$randomforest$call
```

## Pre-trained classifiers

\textbf{Load FFORMS classifier for hourly series}

```r
data("hourly_fforms")
```

```{r, eval=TRUE, size='tiny'}
data("hourly_fforms")
```

\pause

\textbf{Forecast hourly time series in the M4-competition}

```r
fcast.models <- predict(hourly_fforms, features_M4H)
head(fcast.models)
```

```{r, eval=TRUE}
data("features_M4H")
```

```{r, eval=TRUE, size='tiny'}
fcast.models <- predict(hourly_fforms, features_M4H)
head(fcast.models)
```

## \fontsize{10}{10}\sffamily Yearly: Correlation between MASE values across different forecast-models

\centerline{\includegraphics[width=8cm]{img/corY.png}}

## FFORMPP: Feature-based FORecast Model Performance Prediction

\centerline{\includegraphics[width=8cm]{img/mul.png}}

- Efficient Bayesian Multivariate Surface Regression (Feng Li &  Mattias Villani, 2013)
     - handles interactions and nonlinear relationships 
     - allows the knot locations to move freely in the feature space

## `fformpp` R package

\textbf{Installation}

```r
devtools::install_github("thiyangt/fformpp")
library(fformpp)
```

\textbf{Train a model}

\fontsize{11}{11}\sf
```r
fit_fformpp(feamat=features_mat, accmat=forecast.error, 
            sknots=2, aknots=2,
            fix.s=0, fix.a=0, fix.shrinkage=1:5,            
            fix.covariance=0,
            fix.coefficients=0, n.iter=100,
            knot.moving.algorithm="Random-Walk",
            ptype=c("identity", "identity", "identity"),
            prior.knots=100)

```

## FFORMPP: online phase

\fontsize{10}{12}\sf
```r
predict.m1 <- predict(fformpp.model, features.m1.df,
        c("ets", "arima", "rw", "rwd", "wn", "theta", "nn"),
        log=FALSE, final.estimate=median)
head(predict.m1)

```
\fontsize{8}{8}\sf
```{r}
library(fformpp)
data("fformpp.model")
data("features.df.m1")
predict.m1 <- predict_fformpp(fformpp.model, features.df.m1, c("ets", "arima", "rw", "rwd", "wn", "theta", "nn"), log=FALSE, final.estimate=median)
head(predict.m1)
```



## Results: M4 Competition data

\begin{table}[!h]
\centering\scriptsize\tabcolsep=0.12cm
\begin{tabular}{l|rrrrrr}
 & Yearly & Quarterly & Monthly & Weekly & Daily & Hourly \\\hline
\textcolor{red}{FFORMS\_individual} & 3.17 &  1.20 &  0.98&  2.31 & 3.57 &  0.84\\
\textcolor{blue}{FFORMPP\_combination} & 3.07 &  1.13 &  0.89&  2.46 & 3.62 &  0.96\\\hline
auto.arima & 3.40 &1.17  &0.93  & 2.55 &  -& - \\
ets & 3.44 &  1.16& 0.95 &  -&-  &  -\\
theta & 3.37 &1.24  & 0.97 &2.64  & 3.33 & 1.59 \\
rwd & 3.07 & 1.33 & 1.18  & 2.68  & 3.25 & 11.45 \\
rw & 3.97 & 1.48 & 1.21  &2.78  & 3.27 & 11.60 \\
nn & 4.06 & 1.55 &  1.14 &4.04 & 3.90 & 1.09 \\
stlar & - & 2.02 &  1.33& 3.15 & 4.49 & 1.49 \\
snaive & - &  1.66& 1.26 &  2.78& 24.46 & 2.86 \\
tbats & - & 1.19 &  1.05& 2.49 & 3.27 &  1.30\\
wn & 13.42 &  6.50&  4.11&  49.91& 38.07 & 11.68 \\
mstlarima & - & - &  - & - & 3.84 &  1.12\\
mstlets & - &  - &  - &  - & 3.73 &  1.23\\
combination (mean) & 4.09 & 1.58 &  1.16&6.96  & 7.94 & 3.93 \\\hline
M4-1st & 2.98 & 1.12 &  0.88& 2.36 & 3.45 & 0.89\\
M4-2nd & 3.06 & 1.11 &  0.89& 2.11 & 3.34 & 0.81\\
M4-3rd & 3.13 & 1.23 &  0.95& 2.16 & 2.64 & 0.87\\\hline
\end{tabular}
\end{table}

## Thank you

\textbf{R packages and papers}
\fontsize{14}{17}\sf

\begin{alertblock}{R packages}
\begin{itemize}\tightlist
 \item \alert{seer}: FFORMS \newline\textbf{github.com/thiyangt/seer}
 \item \alert{fformpp}: FFORMPP \newline\textbf{github.com/thiyangt/fformpp}
\end{itemize}
\end{alertblock}

\begin{alertblock}{Papers}
Available from \bf{robjhyndman.com}
\end{alertblock}

Slides: \bf{https://thiyanga.netlify.com/talk/user19-talk/}

email: \textcolor{blue}{\bf{thiyanga.talagala@monash.edu}}
