---
title: Feature-based Time Series Forecasting
author: Thiyanga Talagala, \break Rob J Hyndman, George Athanasopoulos \break Feng Li, Yanfei Kang
date:  "11 July 2019"
fontsize: 12pt
classoption: compress
toc: false
output:
  binb::monash:
    fig_height: 5
    fig_width: 8
    highlight: tango
    incremental: no
    keep_tex: no
    includes:
      in_header: preamble.tex
---


```{r setup, include=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE, cache=FALSE, dev.args=list(bg=grey(0.9), pointsize=11))
library(tsfeatures)
library(tidyverse)
library(forecast)
library(Mcomp)
library(pander)
library(seer)
```



## Big picture

\begin{tikzpicture}
    \node<1> (img1) {\centerline{\includegraphics[width=\paperwidth, height=6.3cm]{img/rice1.png}}};
    \node<2> (img2) {\centerline{\includegraphics[width=\paperwidth, height=6.3cm]{img/rice2.png}}};
\end{tikzpicture}

\begin{itemize}[<+->]
\vfill\item What algorithm is likely to perform best?
\vfill\item Algorithm selection problem, John Rice (1976)
\end{itemize}


## Time series features

\fontsize{13}{14}\sf

* Transform a given time series $y=\{y_1, y_2, \cdots, y_n\}$ to a feature vector $F = (f_1(y), f_2(y), \cdots, f_p(y))'$}.

\pause

\begincols
\begincol{.48\textwidth}


```{r, warning=FALSE, message=F, echo=FALSE, fig.height=10}
library(grid)
library(gridExtra)
library(ggrepel)
library(png)
library(tsfeatures)
library(ggpubr)

# Extract required series
series_id <- c("N0001", "N0633", "N0625", "N0645","N1912", "N2012")
six_series <- lapply(M3[series_id], function(u){u$x})
p <- lapply(six_series,
function(u) {autoplot(u) + xlab("") + ylab("")}
)
for (i in seq_along(six_series))
p[[i]] <- p[[i]] + ggtitle(series_id[i])+theme(title =element_text(size=30, face='bold'))
grid.arrange(grobs = p, ncol = 2)

```

\endcol 
\begincol{.48\textwidth}

\pause

```{r fig2, fig.height=6.8, fig.width=7, message=F, echo=F, warning=F}
df <- tsfeatures(six_series, c("stl_features")) %>%
  select(trend, seasonal_strength) %>%
  rename(seasonality = seasonal_strength) %>%
  replace_na(list(seasonality = 0))
df$id <- names(six_series)
ggplot(df, aes(x = trend, y = seasonality)) +
  geom_point(size = 5, color = 'red') +
  xlim(0, 1) + ylim(0, 1) +
  coord_fixed() +
  geom_text_repel(
    aes(label = id),
    colour = "black",
    size = 10,
    box.padding = unit(0.5, "lines")
  ) +
  theme(legend.position = "none", text = element_text(size=20))
```

  \endcol
\endcols

## More features

\begin{textblock}{12}(0.1,1.3)
\begin{multicols}{2}
  \begin{itemize}\tightlist
    \item length
    \item strength of seasonality
    \item strength of trend
    \item linearity
    \item curvature
    \item spikiness
    \item stability
    \item lumpiness
    \item spectral entropy
    \item Hurst exponent
    \item nonlinearity
    \item unit root test statistics
    \item parameter estimates of Holt's linear trend method
    \item parameter estimates of Holt-Winters' additive method
    \item ACF and PACF based features - calculated on raw, differenced, seasonally-differenced series and remainder series.
    \end{itemize}
\end{multicols}
\end{textblock}


## Algorithm selection framework

\begin{tikzpicture}
    \node<1> (img1) {\centerline{\includegraphics[width=\paperwidth]{img/f1.png}}};
    \node<2> (img2) {\centerline{\includegraphics[width=\paperwidth]{img/f2.png}}};
    \node<3> (img3) {\centerline{\includegraphics[width=\paperwidth]{img/f3.png}}};
    \node<4> (img4) {\centerline{\includegraphics[width=\paperwidth]{img/f4.png}}};
    \node<5> (img5) {\centerline{\includegraphics[width=\paperwidth]{img/f5.png}}};
    \node<6> (img6) {\centerline{\includegraphics[width=\paperwidth]{img/f6.png}}};
    \node<7> (img7) {\centerline{\includegraphics[width=\paperwidth]{img/f7.png}}};
    \node<8> (img8) {\centerline{\includegraphics[width=\paperwidth]{img/f8.png}}};
    \node<9> (img9) {\centerline{\includegraphics[width=\paperwidth]{img/f9.png}}};
    \node<10> (img10) {\centerline{\includegraphics[width=\paperwidth]{img/f10.png}}};
    \node<11> (img11) {\centerline{\includegraphics[width=\paperwidth]{img/f11.png}}};
    \node<12> (img12) {\centerline{\includegraphics[width=\paperwidth]{img/f12.png}}};
    \node<13> (img13) {\centerline{\includegraphics[width=\paperwidth]{img/f13.png}}};
    \node<14> (img14) {\centerline{\includegraphics[width=\paperwidth]{img/fh14.png}}};
\end{tikzpicture}

## FFORMS: Feature-based FORecast Model Selection

 \centerline{\includegraphics[width=\paperwidth]{img/fformsd1.png}}
 
* \fontsize{8}{8}{three algorithms: \colorbox{violet}{\color{white}{FFORMS}}, FFORMPP}

## \fontsize{12}{12}\bf\sffamily FFORMPP: Feature-based FORecast Model Performance Prediction

 \centerline{\includegraphics[width=\paperwidth]{img/fformppd1.png}}
 
* \fontsize{8}{8}{three algorithms: \colorbox{violet}{\color{white}{FFORMS}}, \colorbox{cyan}{\color{black}{FFORMPP}}}


## `seer` R package

\textbf{Installation}
```r
devtools::install_github("thiyangt/seer")
library(seer)
```
\begin{tikzpicture}[remember picture,overlay]  
  \node [xshift=-1.5cm,yshift=-2cm] at (current page.north east)
    {\includegraphics[width=\textwidth,height=3cm,keepaspectratio]{img/seer.png}};
\end{tikzpicture}

\pause

\textbf{Example datasets}

\colorbox{blue}{\textbf{\color{white}{observed time series - M1 yearly series (181)}}}

```r
library(Mcomp)
yearlym1 <- subset(M1, "yearly")
```
\pause

```{r echo=FALSE, message=FALSE, warning=FALSE, comment=NA, size='tiny'}
library(Mcomp)
yearlym1 <- subset(M1, "yearly")
```



## Input: features

```r
cal_features(yearlym1[1:3], database="M1",
h=6, highfreq=FALSE)

```
\fontsize{8}{10}\sf


```{r echo=FALSE, message=FALSE, warning=FALSE, comment=NA, size='tiny'}
seer::cal_features(yearlym1[1:2], database="M1", h=6, highfreq=FALSE) 
```

## Input: features

\fontsize{8}{8}\sf


```{r echo=TRUE, message=FALSE, warning=FALSE, comment=NA, size='tiny'}
seer::fcast_accuracy(tslist=yearlym1[1:2], 
               models= c("arima","ets","rw","rwd", "theta", "nn"), 
               database ="M1", cal_MASE, h=6, 
               length_out = 1, 
               fcast_save = TRUE)
```

## Training set

```r
prepare_trainingset(accuracy_set = accuracy_m1, 
feature_set = features_m1)$trainingset
```

\fontsize{8}{10}\sf

```{r echo=FALSE, message=FALSE, warning=FALSE, comment=NA, size='tiny'}
data(M1)
yearly_m1 <- subset(M1, "yearly")
accuracy_m1 <- fcast_accuracy(tslist=yearly_m1[1:2], models= c("arima","ets","rw","rwd", "theta", "nn"), database ="M1", cal_MASE, h=6, length_out = 1, fcast_save = TRUE)
features_m1 <- cal_features(yearly_m1[1:2], database="M1", h=6, highfreq = FALSE)

# prepare training set
prepare_trainingset(accuracy_set = accuracy_m1, feature_set = features_m1)$trainingset
```

## FFORMS classifer

```r
rf <- build_rf(training_set = training_set, 
                testset= M3yearly_features,  rf_type="ru", ntree=100, seed=1, 
               import=FALSE, mtry = 8)
```

\fontsize{8}{10}\sf
```{r echo=FALSE, message=FALSE, warning=FALSE, comment=NA, size='tiny'}
# steps 3 and 4 applied to yearly series of M1 competition
data(M1)
yearly_m1 <- subset(M1, "yearly")
accuracy_m1 <- fcast_accuracy(tslist=yearly_m1, models= c("arima","ets","rw","rwd", "theta", "nn"), database ="M1", cal_MASE, h=6, length_out = 1, fcast_save = TRUE)
features_m1 <- cal_features(yearly_m1, database="M1", h=6, highfreq = FALSE)
# prepare training set
prep_trainingset <- prepare_trainingset(accuracy_set = accuracy_m1, feature_set = features_m1)
training_set <- prep_trainingset$trainingset
yearly_m3 <- subset(M3, "yearly") # M3 dataset is used as a test set
M3yearly_features <- seer::cal_features(yearly_m3, database="M3", h=6, highfreq = FALSE)
rf <- build_rf(training_set = training_set, testset= M3yearly_features,  
               rf_type="ru", ntree=100, seed=1, 
               import=FALSE, mtry = 8)
```

### Predictions
```r
table(rf$predictions)
```

\fontsize{8}{10}\sf
```{r, eval=TRUE, size='tiny'}
head(rf$predictions,10)
```

### FFORMS

\fontsize{8}{10}\sf
```r
rf$randomforest
```
```{r, eval=TRUE, size='tiny'}
rf$randomforest$call
```

## Pre-trained classifiers

### Load hourly data FFORMS classifier

```r
data("hourly_fforms")
```

```{r, eval=TRUE, size='tiny'}
data("hourly_fforms")
```

### Forecast hourly time series in the M4-competition 

```r
fcast.models <- predict(hourly_fforms, features_M4H)
head(fcast.models)
```

```{r, eval=TRUE}
data("features_M4H")
```

```{r, eval=TRUE, size='tiny'}
fcast.models <- predict(hourly_fforms, features_M4H)
head(fcast.models)
```

## FFORMPP: 

